---
title: "Amem Junior Lyrics: A brief word analysis."
author: "by: Mario Saraiva"
date: "Feb, 24, 2020"
output: 
    prettydoc::html_pretty:
    theme: architect
---

##Brief text analysis of Amem Jr (former Salz Band) lyrics. 

29 lyrics (from https://www.letras.mus.br/amen-jr/ and 
https://www.letras.mus.br/salz-band/) were selected for this project:

```{r echo= F, message=FALSE, warning=FALSE}
library("tm")
library("SnowballC")
library("wordcloud")
library("RColorBrewer")
library("prettydoc")
```


```{r echo= F, message=FALSE, warning=FALSE}
library(tm)
corpus.raw <- Corpus(DirSource("/Users/mariosaraiva/Documents/Amem Jr/"))

list.files("/Users/mariosaraiva/Documents/Amem Jr/")
#print(list(DirSource("/Users/mariosaraiva/Documents/Amem Jr/")))
#corpus.raw
```



```{r echo= F, message=FALSE, warning=FALSE}
corpus <- tm_map(corpus.raw, content_transformer(tolower))
corpus <- tm_map(corpus, stripWhitespace) 
corpus <- tm_map(corpus, removePunctuation)
corpus <- tm_map(corpus, removeNumbers)
corpus <- tm_map(corpus, removeWords, stopwords("portuguese"))
# Remove your own stop word
# specify your stopwords as a character vector
corpus <- tm_map(corpus, removeWords, c("pra")) 
corpus <- tm_map(corpus, stemDocument)
```

```{r echo= F, message=FALSE, warning=FALSE}
dtm <- DocumentTermMatrix(corpus) # sparse form
#is.list(dtm)
dtm.mat <- as.matrix(dtm)         # dense form using plain matrices
library(Matrix)                   # sparse form using the Matrix package
dtm.Mat <- sparseMatrix(dtm$i, dtm$j, x = dtm$v, 
                        dims = c(dtm$nrow, dtm$ncol), 
                        dimnames = dtm$dimnames)
```


```{r echo= F, message=FALSE, warning=FALSE}
#Or we could focus on terms that appear at least $1000$ times with
#findFreqTerms(dtm, lowfreq = 40)
```

```{r echo= F, message=FALSE, warning=FALSE}
dtm <- TermDocumentMatrix(corpus)
m <- as.matrix(dtm)
v <- sort(rowSums(m),decreasing=TRUE)
d <- data.frame(word = names(v),freq = v)

```

###Top 20 most frequent words found in the lyrics.
```{r echo= F, message=FALSE, warning=FALSE}
topwords <- head(d, 20)

print(topwords)
```

###Wordcloud of the top 20 words.
```{r echo= F, message=FALSE, warning=FALSE}
set.seed(1234)
wordcloud(words = d$word, freq = d$freq, min.freq = 1,
          max.words = 20, random.order = FALSE, rot.per = 0.35, 
          colors = brewer.pal(8, "Dark2"))
```

###Wordcloud of *all* words.
A quick glance at all the words that appear at least once. 
The words with highest frequency are highlighted in the wordcloud. <p>

```{r echo= F, message=FALSE, warning=FALSE}
set.seed(1234)
wordcloud(words = d$word, freq = d$freq, min.freq = 1,
          max.words = 400, random.order = FALSE, rot.per = 0.35, 
          colors = brewer.pal(12, "Paired"))
```


###Associations between words
We can go a bit deeper into the analysis and look into the association between most frequent words. 
In other words, we are looking at how likely a given word is associated (or closely used with)
another word. 


```{r echo= F, message=FALSE, warning=FALSE}
findAssocs(dtm, "mim", corlimit = 0.5)

findAssocs(dtm, "ser", corlimit = 0.5)

findAssocs(dtm, "tudo", corlimit = 0.5)

findAssocs(dtm, "amor", corlimit = 0.5)

findAssocs(dtm, "vida", corlimit = 0.5)
```

Lastly, we see how the words (that appear at least 10 times) relate to each 
other (with a correlation of at least 0.5).

```{r echo= F, message=FALSE, warning=FALSE}
plot(dtm, terms = findFreqTerms(dtm, lowfreq = 10), 
     corThreshold = 0.5)
```